Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-21 EconGrowth.Rnw
Trying to use apply for energy types and countries. Still not working.

<<loadData>>=
fileName <- "data/USData.txt"

# Read the data file as a table with a header.  
## dataTable is a poor name.  Name it after contents, unless this is the 
## start of a script where this will be provided at the command line.
dataTable <- read.table(fileName, header = TRUE)

# Identifies the header names associated with dataTable
names(dataTable)
@

@@

<<cobb-douglas>>=
# Establish guess values for alpha and lambda.
lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
alphaGuess <- 0.3 # a typical value for alpha, the coefficient on capital stock

@@

# Now create the data for a table.
dataCD <- rbind(upperCD, estCD, lowerCD); dataCD
colnames(dataCD) <- c("$\\lambda$", "$\\alpha$", "$\\beta$", "$\\gamma$")
tableCD <- xtable(dataCD, caption="U.S. Cobb-Douglas, 1980-2011", digit = c(4, 4, 2, 2, 1))
@

@@

<<cobbDouglasEnergy function>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# dataTable is the table to be used for this curve fit
# iEnergy is the name of the column to be used for energy.
#   This may be any of heat (iQ), exergy (iX), or useful work (iU).
##
cobbDouglasEnergy <- function(dataTable, energyColumnName, energyString){
  # Reparameterize to ensure that we meet the constraints:
  # * alpha + beta + gamma = 1.0.
  # * alpha, beta, and gamma are all between 0.0 and 1.0.
  # To do this, we reparameterize as
  # * 0 < a < 1
  # * 0 < b < 1
  # * alpha = min(a, b)
  # * beta = b - a
  # * gamma = 1 - max(a, b)
  iEnergy <- dataTable[energyColumnName] #grabs the energy column
  lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.3 # a typical value for alpha
  betaGuess <- 0.7 # a typical value for beta
  modelCDe <- nls(iGDP ~ exp(lambda*iYear) * 
                    iCapStk^min(a,b) * iLabor^abs(b-a) * 
                    iEnergy^(1.0 - max(a,b)), 
                  algorithm = "port",
                  start = list(lambda=lambdaGuess, a=alphaGuess, b=alphaGuess+betaGuess),
                  lower = list(lambda=-Inf, a=0, b=0),
                  upper = list(lambda=Inf, a=1, b=1),
                  data = dataTable)
  aicCDe <- AIC(modelCDe, k=2) # Checks validity of the model. AIC stands for Akaike's Information Criterion
  print(aicCDe)

@@

<<cdq>>=
dataCDq <- cobbDouglasEnergy(dataTable, "iQ", "q")
# TODO Develop a better caption for the table based on the data automatically.
tableCDq <- xtable(dataCDq, caption="U.S., 1980-2011.", digit = c(4, 4, 2, 2, 3))
#estCDe <- dataCDe[2]
#dataAll <- rbind(estCD, estCDe); dataAll
#tableAll <- xtable(dataAll, caption="U.S. 1980-2011, All Models.", digit = c(4, 4, 2, 2, 3))
@

@@

<<cdx>>=
dataCDx <- cobbDouglasEnergy(dataTable, "iX", "x")
tableCDx <- xtable(dataCDx, caption="U.S., 1980-2011.", digit = c(4, 4, 2, 2, 3))
@

@@

<<cdu>>=
dataCDu <- cobbDouglasEnergy(dataTable, "iU", "U")
tableCDu <- xtable(dataCDu, caption="U.S., 1980-2011.", digit = c(4, 4, 2, 2, 3))
@

>>>>

<<Load Data>>=
#################################
# This function loads data given a country name.
#
# countryName = the name of the country for which data is to be loaded. The file names is assumed to be 
#   named <countryName>Data.txt. The file is assumed to exist in a subfolder of this project called "data".
#
# returns a matrix with the data that has been loaded
##
loadData <- function(countryName){
  # Read the data file as a table with a header.  
  fileName <- paste("data/", countryName, "Data.txt", sep="")
  dataTable <- read.table(fileName, header = TRUE)
  # Identifies the header names associated with dataTable
  names(dataTable)
  return(dataTable)
}
@

@@

<<cobb-douglas>>=

cobbDouglas <- function(countryName){

  # Load the data that we need.
  dataTable <- loadData(countryName)

  # Establish guess values for alpha and lambda.
  lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.3 # a typical value for alpha, the coefficient on capital stock

@@

  # Now create the data for a table.
  dataCD <- rbind(upperCD, estCD, lowerCD); dataCD
  colnames(dataCD) <- c("$\\lambda$", "$\\alpha$", "$\\beta$", "$\\gamma$")
  tableCD <- xtable(dataCD, caption=paste(countryName, " Cobb-Douglas, 1980-2011", sep=""), digit = c(4, 4, 2, 2, 1))
  return(tableCD)
}

tableCD <- cobbDouglas("US")

countries <- c("US", "CN")
tablesCD <- apply(countries, 1, cobbDouglas)
@

@@

<<cobbDouglasEnergy function>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyColumnName is the name of the column to be used for energy.
#   This may be any of heat ("iQ"), exergy ("iX"), or useful work ("iU").
# energyString is the string to be used in table captions reprsenting the type of energy. Typically, "q", "x", or "u"
##
cobbDouglasEnergy <- function(countryName, energyColumnName, energyString){

  # Load the data that we need.
  dataTable <- loadData(countryName)

  # Reparameterize to ensure that we meet the constraints:
  # * alpha + beta + gamma = 1.0.
  # * alpha, beta, and gamma are all between 0.0 and 1.0.
  # To do this, we reparameterize as
  # * 0 < a < 1
  # * 0 < b < 1
  # * alpha = min(a, b)
  # * beta = b - a
  # * gamma = 1 - max(a, b)
  iEnergy <- dataTable[energyColumnName] #grabs the desired energy column
  lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.3 # a typical value for alpha
  betaGuess <- 0.7 # a typical value for beta
  modelCDe <- nls(iGDP ~ exp(lambda*iYear) * 
                    iCapStk^min(a,b) * iLabor^abs(b-a) * 
                    iEnergy^(1.0 - max(a,b)), 
                  algorithm = "port",
                  start = list(lambda=lambdaGuess, a=alphaGuess, b=alphaGuess+betaGuess),
                  lower = list(lambda=-Inf, a=0, b=0),
                  upper = list(lambda=Inf, a=1, b=1),
                  data = dataTable)
  aicCDe <- AIC(modelCDe, k=2) # Checks validity of the model. AIC stands for Akaike's Information Criterion
  print(aicCDe)

@@

<<cdq>>=
countryName <- "US"
energyColumn <- "iQ"; energyString <- "q"
dataCDq <- cobbDouglasEnergy(countryName, energyColumn, energyString)
# TODO Develop a better caption for the table based on the data automatically.
tableCDq <- xtable(dataCDq, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
#estCDe <- dataCDe[2]
#dataAll <- rbind(estCD, estCDe); dataAll
#tableAll <- xtable(dataAll, caption="U.S. 1980-2011, All Models.", digit = c(4, 4, 2, 2, 3))
@

@@

<<cdx>>=
energyColumn <- "iX"; energyString <- "x"
dataCDx <- cobbDouglasEnergy(countryName, energyColumn, energyString)
tableCDx <- xtable(dataCDx, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
@

@@

<<cdu>>=
energyColumn <- "iU"; energySTring <- "u"
dataCDu <- cobbDouglasEnergy(countryName, energyColumn, energyString)
tableCDu <- xtable(dataCDu, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
@
